---
title: "Data Viz-Class"
author: "Desik Somasundaram"
date: "8/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# General use packages:
library(tidyverse)
library(here)
library(janitor)

# Specifically for plots:
library(patchwork)
library(ggrepel)
library(gghighlight)
library(paletteer)
library(ggExtra)
library(ggbeeswarm)

# And for another dataset we'll explore:
library(gapminder)

# Spatial
library(sf)

```

```{r}
lizards <- read_csv(here("data_tidy","lizards.csv"))

```
Histogram (really bad) of lizard weights

Colors- points and lines
Fill- fill

```{r}
ggplot(data = lizards, aes(x = weight)) + 
  geom_histogram(color = "orange", fill = "purple")
```
```{r}
# total length on x-axis
# weight on y-axis

ggplot(data = lizards, aes(x = total_length, y = weight)) + geom_point(shape = 172, fill = "green", color = "blue", size = 4, alpha = 0.5)
```

 
```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) + 
  geom_point(aes(color = common_name)) + 
  theme_light() +
  facet_wrap(~site, ncol = 3)
```
```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) + 
  geom_point(aes(color = common_name)) + 
  theme_light() +
  facet_grid(sex ~ tail)
```

```{r}
lizard_counts <- lizards %>% 
  group_by(common_name) %>% 
  summarize(count = n())


```

```{r}
ggplot(data = lizard_counts, aes(y = fct_reorder(common_name, count), x = count)) +
  geom_col(aes(fill = common_name)) +
  labs( x = "Lizard counts", y = "Common Names")

```
```{r}
lizards_reorder <- lizard_counts %>% 
  mutate(common_name = fct_reorder(common_name, count))
```

## Axis scale break, limit & labels

Scatterplot: total_length (x) versus weight (y)


```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) + 
  geom_point() +
  scale_x_continuous(breaks = c(50, 100, 200, 500),
                     limits = c(0,500)) 
```
```{r}
lizard_counts <- lizards %>%
  mutate(date = lubridate::mdy(date)) %>%
  group_by(date) %>% 
  summarize(count = n())

class(lizard_counts$date)

```
```{r}
ggplot(data = lizard_counts, aes(x = date, y = count)) +
  geom_line() +
  scale_x_date(breaks = scales::breaks_width("3 years"), labels = scales::label_date("%y"))
```
```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point((aes(color = weight))) +
  scale_color_gradientn(colors = c("red", "navy", "cyan", "chartreuse", "orange"))

```
```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) +
  scale_color_steps2(colors = c("green", "blue", "purple"),
                     breaks = c(0, 20, 40, 50, 100))
```
```{r}
lizard_fct <- lizards %>% 
  mutate(common_name = fct_reorder(common_name, total_length, .fun = median))


ggplot(data = lizard_fct, aes(x = total_length, y = common_name)) + 
  geom_boxplot(aes(fill = common_name), show.legend = FALSE) +
  scale_fill_paletteer_d("nbapalettes::nuggets_80s")
```
```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  theme(
    panel.background = element_rect(fill = "yellow", 
                                    color = "purple",
                                    size = 10),
    panel.grid.major.x = element_line(color = "blue"),
    panel.grid.major.y = element_line(color = "green"),
    plot.background = element_rect(fill = "cyan")
  )
```

## ggrepel

```{r}

ww_lizards_1 <- lizards %>% 
  filter(common_name == "western whiptail",
         site == "sand")

ggplot(data = WW_lizards_1, aes(x = total_length, y = weight)) + 
  geom_point() +
  geom_text_repel(aes(label = toe_num), size = 3, max.overlaps = 20)

```
```{r}

gapminder_subset <- gapminder %>% 
  filter(continent == "Europe",
         year == "2007")

ggplot(data = gapminder_subset, aes(x = gdpPercap, y = lifeExp)) +
  geom_point() +
  geom_text_repel(aes(label = country), size = 3)
```
```{r}
p <- ggplot(lizards, aes(x = total_length, y = weight)) +
  geom_point()
    
p + gghighlight(toe_num == 250, label_key = toe_num)
```
```{r}
q <- ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_line(aes(color = common_name)) +
  gghighlight(max(weight) > 30)


q 
```
## patchwork for compound figures

```{r}

p + q &
  theme_minimal()



```
## A few graph types

### Marginal plots

```{r}
whiptails <- lizards %>% 
  filter(common_name == "western whiptail") %>% 
  drop_na(total_length, weight)

ggplot(data = whiptails, aes(x = total_length, y = weight)) +
  geom_point() +
  geom_rug()


```
```{r}
myplot<- ggplot(data = whiptails, aes(x = total_length, y = weight)) +
  geom_point(aes(color = sex), size = 2) +
  scale_color_manual(values = c("cyan4", "black", "goldenrod"),
                     name = "Sex:",
                     labels = c("female", "juvenile", "male")) +
  theme_minimal() +
  theme(legend.position = "bottom")

ggMarginal(myplot, type = "boxplot", groupColour = TRUE)
```
### Beeswarm plot

```{r}
ggplot(data = whiptails, aes(x = sex, y = weight)) + 
  geom_beeswarm() +
  geom_boxplot(fill = NA)
```
### A heatmap with geom_tile()

Make a new data frame called lizard_count, starting from lizards, with data converted to class date. 

```{r}
lizard_count <- lizards %>%
  mutate(date = lubridate::mdy(date),
         year = lubridate::year(date)) %>% 
  group_by(year, common_name) %>%
  summarize(total = n())
  
```

```{r}
ggplot(data = lizard_count, aes(x = year, y = common_name)) +
  geom_tile(aes(fill = total)) +
  geom_text(aes(label = total), color = "white", size = 3) +
  scale_fill_gradientn(colors = c("navy", "blue", "purple"))

```
### Make a map!

```{r}
jornada_vegetation <- read_sf(here("data_raw", "spatial_vegetation", "doc.kml")) %>%
  select(Name) %>% 
  clean_names()

ggplot(jornada_vegetation) + 
  geom_sf(aes(fill = name), color = NA) +
  scale_fill_paletteer_d(palette = "ggthemes::manyeys") +
  labs(x = "Longitude",
       y = "Latitude",
       fill = "Dominant vegeation:")

```

